\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{xcolor}

\onehalfspacing

\title{\textbf{Experimental Documentation}\\
Multi-Task CNN for Mammogram Mass ROI Classification}
\author{Ali Saeidi Rad}
\date{27 February 2026}

\begin{document}

\maketitle

\begin{abstract}
This report documents the experimental design, configurations, and outcomes of a multi-task convolutional neural network (CNN) developed for classification of mammographic mass regions of interest (ROIs). The network simultaneously predicts morphological and clinical attributes, including shape, margin, BI-RADS assessment, and pathology. Multiple training strategies, preprocessing pipelines, and dataset combinations were evaluated to determine an optimal configuration.
\end{abstract}

\section{Experiment 1}

\subsection*{Overview}

In the initial experiment, the network consisted of four prediction heads:
\textit{Shape}, \textit{Margin}, \textit{BI-RADS}, and \textit{Pathology}.

\subsection*{Configuration}

\begin{enumerate}
    \item \textbf{Backbone:} \texttt{convnext\_base.fb\_in22k\_ft\_in1k}
    
    \item \textbf{Optimizer:} \texttt{AdamW}
    
    \item \textbf{Scheduler:} \texttt{CosineAnnealingWarmRestarts}
    \begin{enumerate}
        \item Initial learning rate: $\eta_{0} = 10^{-5}$
        \item Minimum learning rate: $\eta_{\min} = 10^{-8}$
        \item Initial restart period: $T_{0} = 20$
        \item Multiplication factor: $T_{\text{mult}} = 2$
    \end{enumerate}
    
    \item \textbf{Image resizing backbone:} OpenCV
    
    \item \textbf{Preprocess:} Normal
    
    \item \textbf{Datasets:} CBIS-DDSM + VIDA
\end{enumerate}

\subsection*{Results}

\begin{table}[h]
\centering
\small
\caption{Performance comparison on the test dataset with and without weighted TTA.}
\begin{tabular}{lcccccc}
\toprule
& \multicolumn{3}{c}{\textbf{Test Dataset}} 
& \multicolumn{3}{c}{\textbf{Test Dataset + Weighted TTA}} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
\textbf{Head} 
& F1 (\%) & Acc (\%) & AUC (\%) 
& F1 (\%) & Acc (\%) & AUC (\%) \\
\midrule
Margin    & 49.87 & 46.79 & 82.99 & 47.95 & 46.15 & 82.13 \\
Shape     & 41.49 & 39.74 & 73.10 & 41.90 & 39.74 & 73.77 \\
BI-RADS   & 48.11 & 50.00 & 84.20 & 51.62 & 53.85 & 83.77 \\
Pathology & 73.90 & 73.72 & 83.75 & 74.54 & 74.36 & 84.82 \\
\bottomrule
\end{tabular}
\end{table}

\section{Experiment 10 (Nominated Strategy)}
After extensive hyperparameter tuning and pipeline modifications, the following configuration was selected as the most stable and effective.

\subsection*{Configuration}

\begin{enumerate}
    \item \textbf{Backbone:} \texttt{convnext\_base.fb\_in22k\_ft\_in1k}
    
    \item \textbf{Optimizer:} \texttt{AdamW}
    
    \item \textbf{Scheduler:} \texttt{CosineAnnealingWarmRestarts}
    \begin{enumerate}
        \item Initial learning rate: $\eta_{0} = 10^{-5}$
        \item Minimum learning rate: $\eta_{\min} = 10^{-8}$
        \item Initial restart period: $T_{0} = 20$
        \item Multiplication factor: $T_{\text{mult}} = 2$
    \end{enumerate}
    
    \item \textbf{Image resizing backbone:} \textcolor{red}{Pillow}
    
    \item \textbf{Preprocess:} \textcolor{red}{ISS}
    
    \item \textbf{Datasets:} CBIS-DDSM + VIDA
\end{enumerate}

\subsection*{Observed Effects of Design Choices} The following findings were consistently observed across experiments:
in conducted experiments we find out that following changes has no effect or have negative effects:

\begin{enumerate}
    \item OpenCV-based resizing degraded performance, likely due to excessive loss of fine structural details. 
    \item Standard normalization alone was insufficient for robust learning. 
    \item Removing any prediction head reduced overall performance, indicating beneficial multi-task feature sharing. 
    \item BI-RADS predictions alone were not reliable for inferring pathology outcomes. 
\end{enumerate}

\begin{table}[h]
\centering
\small
\caption{Performance comparison on the test dataset with and without weighted TTA.}
\begin{tabular}{lcccccc}
\toprule
& \multicolumn{3}{c}{\textbf{Test Dataset}} 
& \multicolumn{3}{c}{\textbf{Test Dataset + Weighted TTA}} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
\textbf{Head} 
& F1 (\%) & Acc (\%) & AUC (\%) 
& F1 (\%) & Acc (\%) & AUC (\%) \\
\midrule
Margin    & 51.11 & 48.07 & 82.91 & 51.85 & 50.00 & 83.20 \\
Shape     & 46.66 & 46.15 & 74.34 & 45.54 & 45.51 & 74.56 \\
BI-RADS   & 51.27 & 52.56 & 83.45 & 51.70 & 52.56 & 84.37 \\
Pathology & 73.20 & 73.07 & 85.10 & 71.89 & 71.79 & 85.62 \\
\bottomrule
\end{tabular}
\end{table}

\section{Experiment 24 (Final Pipeline)}
Subsequent experiments incorporated additional datasets and an extra prediction head for malignancy probability when pathology labels were unavailable.

\subsection*{Training Strategy}

\begin{enumerate}
   \item \textbf{Stage 1 — Pretraining:} Train on CBIS-DDSM and CDD-CESM for 700 epochs with frozen backbone using $lr_{0}=10^{-5}$.
   \item \textbf{Stage 2 — Full Fine-Tuning:} Continue training for 700 epochs with backbone unfrozen using a reduced learning rate $lr_{0}=10^{-6}$.
    \item \textbf{Stage 3 — Dataset-Specific Fine-Tuning:} Fine-tune individual heads (maximum 300 epochs, early stopping patience of 30 epochs):
    \begin{enumerate}
        \item VIDA: Train Pathology head only; freeze attention modules and backbone.
        \item CSAW: \textcolor{red}{Excluded due to unreliable annotations.}
        \item VINDR: Train relevant heads while freezing Pathology, Margin, Shape, attention modules, and backbone.
    \end{enumerate}
\end{enumerate}

\begin{table}[h]
\centering
\small
\caption{\textcolor{red}{UNDER GOING}}
\begin{tabular}{lcccccc}
\toprule
& \multicolumn{3}{c}{\textbf{Test Dataset}} 
& \multicolumn{3}{c}{\textbf{Test Dataset + Weighted TTA}} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
\textbf{Head} 
& F1 (\%) & Acc (\%) & AUC (\%) 
& F1 (\%) & Acc (\%) & AUC (\%) \\
\midrule
Margin    & - & - & - & - & - & - \\
Shape     & - & - & - & - & - & - \\
BI-RADS   & - & - & - & - & - & - \\
Pathology & - & - & - & - & - & - \\
\bottomrule
\end{tabular}
\end{table}

\end{document}